% Filled (batch 2): gate ⑥ Lawson 0-D power balance derivation.
% Source: stdlib/fusion/lawson.* — triple_product / q_plasma / ignition_margin
% atoms (verify-ledger.json, PR #1012, gate (b)). Pure rational, no libm.
\section{Lawson 0-D Power Balance}
\label{app:lawson}

This appendix is the full data record for pipeline stage
\textcircled{\scriptsize 6} (gate b): the 0-D power-balance derivation of
the Lawson triple product $nT\tau_{\!E}$ and its D--T ignition threshold,
the 0-D plasma gain $Q_{\text{plasma}}=P_{\text{fus}}/P_{\text{heat}}$, and
the ignition margin. Every kernel is a pure rational expression (no libm),
so each intermediate is exact and carries a \tierGreen\ verify atom at
$|\Delta|=0$. Data source: the \verb|triple_product|, \verb|q_plasma| and
\verb|ignition_margin| atoms in \texttt{companion/verify-ledger.json}
(PR~\#1012, gate~(b)).

% -------------------------------------------------------------------------
\subsection{0-D power balance and the ignition criterion}
\label{app:lawson-balance}

Treat the plasma as a single zone with density $n$, temperature $T$, stored
thermal energy $W=3nT V$ (volume $V$), and energy-confinement time
$\tau_{\!E}$, defined operationally by the loss rate $P_{\text{loss}}=W/\tau_{\!E}$.
The 0-D power balance in steady state is
\begin{equation}
\underbrace{P_\alpha}_{\text{$\alpha$ self-heat}}
  \;+\;
\underbrace{P_{\text{heat}}}_{\text{external}}
  \;=\;
\underbrace{\frac{W}{\tau_{\!E}}}_{\text{transport loss}},
\qquad
P_\alpha \;=\; f_\alpha\,P_{\text{fus}}
  \;=\;
  \tfrac14\,n^2\langle\sigma v\rangle\,E_\alpha\,V,
\label{eq:balance}
\end{equation}
where for D--T at $50$:$50$ fuel the $\alpha$ carries $f_\alpha=E_\alpha/
E_{\text{fus}}=3.52/17.6=0.2$ of the fusion power and the neutron carries the
rest. \emph{Ignition} is the point at which the $\alpha$ self-heat alone
sustains the balance, i.e. $P_{\text{heat}}\to 0$:
\begin{equation}
f_\alpha\,P_{\text{fus}} \;\ge\; \frac{W}{\tau_{\!E}}
\;\;\Longleftrightarrow\;\;
\tfrac14 n^2\langle\sigma v\rangle E_\alpha V \;\ge\; \frac{3nTV}{\tau_{\!E}}.
\label{eq:ignite}
\end{equation}
Rearranging \eqref{eq:ignite} and folding the temperature dependence of
$\langle\sigma v\rangle/T^2$ (which is near-constant over the D--T burn
window $\sim\!10$--$20$\,keV) gives the canonical Lawson triple-product
threshold
\begin{equation}
n\,T\,\tau_{\!E}
  \;=\;
  \frac{12\,T^2}{E_\alpha\,\langle\sigma v\rangle}
  \;\gtrsim\;
  3\times10^{21}\ \text{keV}\,\text{s}\,\text{m}^{-3}
  \qquad(\text{D--T ignition}).
\label{eq:triple}
\end{equation}

\paragraph{Kernel and verify atom.}
The stdlib kernel evaluates the triple product directly as the rational
product $nT\tau_{\!E}$ (the discriminating value, not the threshold
inequality). For the reference point $(n,T,\tau_{\!E})=(10^{20}\,\text{m}^{-3},
15\,\text{keV},3\,\text{s})$:
\begin{equation}
nT\tau_{\!E} \;=\; 10^{20}\cdot 15\cdot 3 \;=\; 4.5\times10^{21}\ \text{keV}\,\text{s}\,\text{m}^{-3},
\end{equation}
which clears the $3\times10^{21}$ D--T threshold with a $1.5\times$ margin.
\begin{lstlisting}
fn: triple_product  args:[1.0e20, 15.0, 3.0]  exp:4.5e21  obs:4.5e21
                    |delta|:0.0  tier:SUPPORTED-NUMERICAL  gate:(b)  #1012
                    note: Lawson triple product nTtau; pure rational, no libm
\end{lstlisting}

% -------------------------------------------------------------------------
\subsection{0-D plasma gain $Q$}
\label{app:lawson-q}

The fusion gain is the ratio of fusion power produced to external heating
power injected,
\begin{equation}
Q_{\text{plasma}} \;=\; \frac{P_{\text{fus}}}{P_{\text{heat}}},
\label{eq:q}
\end{equation}
so $Q=1$ is scientific break-even (fusion power equals heating power),
$Q=5$ is the ITER target, and $Q\to\infty$ is ignition
($P_{\text{heat}}\to0$ in Eq.~\ref{eq:balance}). The kernel is the bare
ratio; for $(P_{\text{fus}},P_{\text{heat}})=(50,10)$\,MW:
\begin{equation}
Q_{\text{plasma}} \;=\; 50/10 \;=\; 5.0.
\end{equation}
\begin{lstlisting}
fn: q_plasma  args:[50.0, 10.0]  exp:5.0  obs:5.0  |delta|:0.0
              tier:SUPPORTED-NUMERICAL  gate:(b)  #1012   note: Q = P_fus / P_heat
\end{lstlisting}

% -------------------------------------------------------------------------
\subsection{Ignition margin}
\label{app:lawson-margin}

The ignition margin compares the $\alpha$ self-heating against the transport
loss directly (the left vs right side of Eq.~\ref{eq:ignite}):
\begin{equation}
\text{ignition margin}
  \;=\;
  \frac{f_\alpha\,P_{\text{fus}}}{W/\tau_{\!E}}
  \;=\;
  \frac{f_\alpha\,P_{\text{fus}}\,\tau_{\!E}}{W},
\label{eq:margin}
\end{equation}
so margin $\ge 1$ is ignited ($\alpha$ heat alone sustains the balance) and
margin $<1$ requires external drive. For
$(P_{\text{fus}},W,\tau_{\!E},f_\alpha)=(50\,\text{MW},100\,\text{MJ},
3\,\text{s},0.2)$:
\begin{equation}
\text{margin} \;=\; \frac{0.2\cdot 50\cdot 3}{100} \;=\; \frac{30}{100} \;=\; 0.3,
\end{equation}
i.e. the $\alpha$ self-heat supplies $30\%$ of the loss at this point ---
sub-ignition, as expected for a $Q=5$ operating point.
\begin{lstlisting}
fn: ignition_margin  args:[50.0, 100.0, 3.0, 0.2]  exp:0.3  obs:0.3  |delta|:0.0
                     tier:SUPPORTED-NUMERICAL  gate:(b)  #1012
                     note: f_alpha*P_fus/(W/tau)
\end{lstlisting}

This margin is the (b) factor of the F-funnel composite (\S\ref{sec:funnel}):
the $\mathrm{clamp}_{[0,1]}$ of \eqref{eq:margin} feeds the $\text{(a)}\times
\text{(b)}$ rank, and the \verb|f_margin_coverage(0.3,0.3)=1.0| /
\verb|f_funnel_composite| atoms (PR~\#1020) confirm the clamp passes the
margin through unchanged at the top-ranked $T=20$\,keV point.

% -------------------------------------------------------------------------
\subsection{Honest scope --- $\tau_{\!E}$ is an input, not a prediction}
\label{app:lawson-honest}

The three kernels above are \emph{pure rational expressions with no libm
calls}, so every intermediate is exact and each atom reports
$|\Delta|=0$ at $\varepsilon=10^{-9}$ --- there is no floating-point
transcendental to lose precision. That exactness is precisely why the honest
scope must be stated plainly: \textbf{$\tau_{\!E}$ is an assumed input, not a
transport prediction}. None of \eqref{eq:triple}, \eqref{eq:q} or
\eqref{eq:margin} \emph{computes} the confinement time; $\tau_{\!E}=3$\,s is
supplied as a parameter (it would, on a real device, come from an empirical
confinement scaling such as IPB98(y,2) or from a turbulent-transport
solver, neither of which this 0-D balance attempts). The gate therefore
certifies the \emph{power-balance arithmetic} given $(n,T,\tau_{\!E})$, and
nothing about whether a given machine \emph{achieves} that $\tau_{\!E}$. The
records carry \texttt{absorbed=false}: clearing gate~(b) is a witness that a
self-consistent 0-D energy balance exists at the stated point, downstream of
which transport-predicted $\tau_{\!E}$ on a built, heated plasma (ultimately
the gate~(g) burning-plasma measurement) remains the wet-lab confirmation.
