\section{Wrap-and-Fix Adapter Defect Catalog}
\label{app:adapters}

This appendix is the full data record for the three wrap-and-fix
defects summarized in Table~\ref{tab:defects}
(\S\ref{sec:defects}). It transcribes, verbatim, the
\texttt{companion/\allowbreak adapter-defect-catalog.json} record for each
defect --- the OpenMC transport-less gate-classifier stub
(FUSION-DEFECT-001), the FreeGS hardcoded coil-set divergence with
\texttt{stdout}-corrupted JSON (FUSION-DEFECT-002), and the GetDP
\texttt{shutil.which}-only checker with the mis-diagnosed $\sim$1.4\%
residual (FUSION-DEFECT-003) --- carrying the origin commit, the
symptom, the root cause, the fix PR, the fix summary, the tier after
fix, and the honesty note.

\subsection{The wrap-and-fix-as-integration-test thesis}
\label{app:adapters:thesis}

The architecture admits a backend in two stages. First, an adapter
\emph{wraps} the external solver behind a uniform contract: it declares
the backend's presence, builds the problem, invokes the solve, and emits
a single machine-readable verdict line. Second, when the backend is
actually installed and the adapter is run end-to-end, any gap between
the wrap and the real solve surfaces as a runtime failure. The thesis of
this appendix is the empirical observation that drove the whole
promotion round:

\begin{quote}
\emph{Every one of the three wrap-and-fix runs surfaced a latent defect
in the landed adapter. The simple act of installing the backend and
running the adapter end-to-end turned the wrap layer into an integration
test.}
\end{quote}

This is not incidental. An install-gated ``honest-skip'' adapter ---
one that reports PASS merely because the backend \emph{could} be
imported --- is a classifier of the \emph{environment}, not of the
\emph{physics}. It can be green for years without ever solving a
problem. The forcing function is to run it for real: the moment a true
solve is demanded, a transport-less stub cannot produce a tally, a
divergent coil set cannot find a magnetic axis, and a binary-presence
check cannot return a field value. The defect is exposed not by code
review but by execution. All three fixes preserve
\texttt{absorbed=false}: a computed witness (a TBR number, a converged
equilibrium, an on-axis $B_z$) is a simulation PASS, not a built machine.

\subsection{FUSION-DEFECT-001 --- OpenMC transport-less stub}
\label{app:adapters:openmc}

\begin{description}
  \item[\texttt{file}] \texttt{stdlib/fusion/openmc\_tbr.py}
  \item[\texttt{origin\_commit}] \texttt{a7a185b} (origin/main)
  \item[\texttt{fix\_pr}] \#1078
  \item[\texttt{tier\_after}] \tierGreen\ \textsc{Supported-Numerical}
\end{description}

\paragraph{Symptom.}
The landed adapter imported OpenMC and checked the ENDF data
environment variable, but \emph{never built a geometry and never invoked}
\verb|openmc.run|. It reported a fake install-gated PASS: green if the
package was importable, regardless of whether any transport had been
solved.

\paragraph{Root cause.}
PR~\#1032-era scaffolding committed the wrap as a \verb|.py|
syntax-check without a transport block. Nobody flagged it because
``check OpenMC importable'' was the gate's first-pass exit criterion ---
the adapter satisfied its own (too-weak) contract. The gate was
classifying the \emph{environment}, not the \emph{physics}.

\paragraph{Fix.}
A new \verb|.hexa| adapter (\verb|stdlib/fusion/openmc_tbr.hexa|) builds
the spherical Pb-15.7Li breeder geometry (W armor / EUROFER first wall /
SS316 shield), runs a $5\times10^{6}$-history Monte Carlo transport via
\verb|openmc.run|, and parses the TBR tally directly. The validated
result is $\text{TBR} = 1.3222 \pm 0.0004$ (the $\pm0.0004$ is the
$1\sigma$ MC statistical uncertainty). The \verb|.py| stub is retained
as a historical artifact but is no longer on the call path. Full
derivation in Appendix~\ref{app:tbr}.

\paragraph{Honesty note.}
\texttt{absorbed=false} preserved --- the TBR is a simulation PASS only;
the ENDF/B-VIII.0 corpus is the data-bound dependency, and a measured
TBR remains downstream wet-lab confirmation.

\subsection{FUSION-DEFECT-002 --- FreeGS coil divergence + stdout JSON corruption}
\label{app:adapters:freegs}

\begin{description}
  \item[\texttt{file}] \texttt{stdlib/fusion/f3\_grad\_shafranov\_adapter.hexa}
  \item[\texttt{origin\_commit}] PR~\#1032
  \item[\texttt{fix\_pr}] \#1075
  \item[\texttt{tier\_after}] \tierGreen\ \textsc{Supported-Numerical}
\end{description}

\paragraph{Symptom.}
The Picard solve diverged with ``\texttt{No O points found}'' on a
custom hardcoded coil set: FreeGS\,0.8.2 could not locate a magnetic
axis. Worse, the warning was emitted on \emph{stdout}, which corrupted
the adapter's JSON line --- so even when the solve technically completed,
there was no machine-readable output for the gate to consume.

\paragraph{Root cause --- two stacked defects.}
\begin{enumerate}
  \item \textbf{Geometric inconsistency.} The hardcoded custom coil
        currents had no admissible equilibrium for that coil-current
        pair --- there was no magnetic configuration the Grad--Shafranov
        operator could converge to.
  \item \textbf{Stream confusion.} FreeGS's early-Picard diagnostic was
        sent to \texttt{stdout}, not \texttt{stderr}, and the adapter
        scraped \texttt{stdout} for the JSON tail. Even a converging run
        would have its output polluted by the warning text.
\end{enumerate}
The two defects masked each other: the divergence hid the stream bug,
and the corrupted output made the divergence hard to diagnose from the
gate's perspective.

\paragraph{Fix.}
Switched to the canonical \verb|freegs.machine.TestTokamak()| preset,
used the \verb|freeBoundaryHagenow| boundary-integral method, and raised
the X-point isoflux constraint to converge cleanly. The solve block was
wrapped in a \texttt{stderr}-redirect context so that the (now spurious)
``No O points'' warning does not pollute \texttt{stdout}. Result:
$I_p = 200$\,kA, $q_0 = 1.355$, $q_{95}\approx 10.08$,
$\kappa = 1.358$, magnetic axis at $(R, Z) = (1.280,\,0.038)$\,m. Full
convergence and $q$-profile detail in Appendix~\ref{app:confinement}.

\paragraph{Honesty note.}
\texttt{absorbed=false} --- the converged equilibrium is the design
witness, not a built machine.

\paragraph{Operational coda.}
This was the first wrap-and-fix run in which an external solver was
actually executed on free pool hardware, and it is the run that
established the integration-test thesis. The cross-repo PR-landing
mechanics of this fix (a demiurge worktree creating a hexa-lang PR,
\texttt{gh} mis-resolving the default repo) are documented as a war
story in Appendix~\ref{app:operational}.

\subsection{FUSION-DEFECT-003 --- GetDP no-solve stub + the false $2\pi$ diagnosis}
\label{app:adapters:getdp}

\begin{description}
  \item[\texttt{file}] \texttt{stdlib/rtsc/getdp\_hts.py}
  \item[\texttt{origin\_commit}] pre-PR~\#1042
  \item[\texttt{fix\_pr}] \#1095
  \item[\texttt{tier\_after}] \tierGreen\ \textsc{Supported-Numerical}
        (FEM bridge) $+$ \tierBlue\ \textsc{Supported-Formal} (closed-form
        anchor)
\end{description}

\paragraph{Symptom.}
A \verb|shutil.which("getdp")| checker only: the adapter returned
``install-gated'' if the GetDP binary was present, but never invoked it
to solve a problem. The \texttt{magnet\_gate} fell back to the
closed-form $1/R$ baseline indefinitely, never exercising the FEM path.
When the FEM path was finally exercised, the on-axis $B_z$ disagreed
with the textbook value by $\sim1.4\%$ --- and that residual was
\emph{initially mis-diagnosed} as a $2\pi$/VolAxiSqu units bug in the
\verb|.pro| deck Jacobian.

\paragraph{Root cause --- two layered issues.}
\begin{enumerate}
  \item \textbf{No-solve stub.} The \verb|.py| wrap-and-fix layer never
        reached the solve invocation; the binary-presence check was the
        \emph{entire} adapter.
  \item \textbf{Far-field truncation (the real fix).} When the actual
        GetDP solve was eventually run by hand, the on-axis $B_z$
        disagreed with the textbook by $\sim1.4\%$. Six hours of
        investigation chased a hypothesized $2\pi$ Jacobian /
        \texttt{VolAxiSqu} units bug in the \verb|.pro| deck. The
        diagnosis was \emph{wrong}. The real cause was far-field
        truncation: the air-region outer radius/height was only
        $\sim5\times$ the coil radius, leaking flux. Pushing
        $R_{\text{out}}/H_{\text{out}}$ to $\sim10\times$ collapsed the
        residual to $<0.1\%$ \emph{with no deck change} --- same
        Jacobian, same \texttt{VolAxiSqu}, same exact deck. The
        truncation-vs-residual sweep is tabulated in
        Appendix~\ref{app:magnet}, \S\ref{app:magnet:farfield}.
\end{enumerate}

\paragraph{Fix.}
Wrote \verb|stdlib/rtsc/getdp_solenoid_solve.hexa| as a hexa-native
driver: it builds the validated \verb|.pro| deck with $10\times$
far-field truncation, runs GetDP (3.5.0 on the Linux pool, 4.0.0 on
Apple silicon as a cross-platform check), parses the on-axis $B_z$ token,
and emits a single \verb|B_peak -> <value>| line on \texttt{stdout}. The
\verb|magnet_gate.hexa| now greps that token and flips
\verb|field_path = "getdp"|. Parity with the closed-form anchor
(\verb|solenoid_axis_bz|, Appendix~\ref{app:verify-ledger}) is
$\Delta = -0.064\%$ at a $\sim33\,$k-node mesh --- well inside the
engineering ``few-percent'' band.

\paragraph{Honesty note --- the recorded false diagnosis.}
The $2\pi$/VolAxiSqu diagnosis was a controlled-falsification artifact.
We record it \emph{explicitly} in the defect catalog --- the wrong
hypothesis, not just the right answer --- to spare the next implementer
the same six-hour chase. The discipline here is the same one that keeps
\tierRed\ negative controls in the verify ledger: a falsified hypothesis
is evidence, and deleting it would erase the most expensive lesson of the
session. \texttt{absorbed=false}: a $1.48$\,T on-axis $B_z$ is a computed
witness, not a built TF coil.

\subsection{Why the false diagnosis is the load-bearing entry}
\label{app:adapters:loadbearing}

Of the three defects, FUSION-DEFECT-003 is the one this catalog exists to
preserve. The first two are stub-replacement stories with clean
before/after states: a transport-less classifier becomes a real
transport run; a divergent coil set becomes a canonical converging
preset. Their lessons are real but local.

The GetDP residual is different because the \emph{obvious} explanation
was wrong in an instructive way. A $1.4\%$ error in an axisymmetric
$A_\phi$ FEM, in the presence of a $2\pi$ azimuthal-integration factor
and a \texttt{VolAxiSqu} Jacobian, \emph{looks exactly like} a units bug
--- $2\pi \approx 6.28$, and a single missing or doubled azimuthal
factor would produce an $\mathcal{O}(1)$ error, but a subtle Jacobian
mis-scaling can equally produce a small few-percent residual. The
hypothesis was plausible enough to consume six hours. The actual cause
--- a finite air region truncating the far field and leaking return flux
--- is invisible in the deck's algebra and visible only in the
\emph{geometry}: it is fixed by enlarging $R_{\text{out}}/H_{\text{out}}$
from $\sim5\times$ to $\sim10\times$ the coil radius, touching no
equation. The diagnostic reversal --- from ``the math is wrong'' to
``the domain is too small'' --- is the single most transferable finding
of the promotion round, which is why the wrong diagnosis is written down
on purpose. The operational framing of this reversal as a codegen /
diagnosis pitfall is carried in Appendix~\ref{app:operational}.

\subsection{Catalog summary and provenance}
\label{app:adapters:summary}

\begin{table}[h]
\centering
\scriptsize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}llll@{}}
\toprule
id & file & PR & tier after \\
\midrule
FUSION-DEFECT-001 & \texttt{fusion/openmc\_tbr.py} & \#1078 & \tierGreen \\
FUSION-DEFECT-002 & \texttt{fusion/f3\_grad\_shafranov\_adapter.hexa} & \#1075 & \tierGreen \\
FUSION-DEFECT-003 & \texttt{rtsc/getdp\_hts.py} & \#1095 & \tierGreen\,\tierBlue \\
\bottomrule
\end{tabular}
\caption{The three wrap-and-fix defects, all surfaced by the act of
running the backend end-to-end, all fixed with
\texttt{absorbed=false} preserved. Source:
\texttt{companion/adapter-defect-catalog.json}.}
\label{tab:app:adapters}
\end{table}

All three defects share the same shape: a wrap layer that satisfied a
too-weak contract (importable / present / syntactically valid) until a
real solve was demanded, at which point the gap became a runtime
failure. The fix in each case was to replace the stub with a hexa-native
adapter that performs the actual solve and emits a machine-readable
witness. The catalog --- \texttt{companion/adapter-defect-catalog.json},
of which this appendix is the verbatim rendering --- is kept so that a
downstream reader does not re-walk the same chases, and so that the one
deliberately-recorded \emph{wrong} diagnosis (FUSION-DEFECT-003) stays on
the record.
