\section{Operational Findings}
\label{app:operational}

This appendix is the full data record for the session's operational
lessons, drawn from \texttt{companion/\allowbreak session-journal.md} ---
the time-ordered narrative of what was changed, what was verified, and
what was left open. Where Appendices~\ref{app:verify-ledger}
and~\ref{app:adapters} record the \emph{physics} witnesses, this
appendix records the \emph{meta-process}: the failure modes of running
many verification agents in parallel against a shared code home, the
codegen-precision pitfalls that masqueraded as physics bugs, and the
worktree-concurrency contract (@D~d9) that kept the shared index clean.
Each finding is stated as a symptom, the diagnosis, and the standing
rule it became.

\subsection{Sustained-parallelism failure: rate-limit agent death and the salvage pattern}
\label{app:operational:ratelimit}

\paragraph{Symptom.}
At one point in the session, a round fanned out $4{+}$ concurrent
verification agents instead of holding the $2$--$3$ cap. Within minutes,
one agent was killed by a provider rate limit mid-task. The killed
agent's in-flight work vanished --- there was no \texttt{git stash}, no
saved buffer, nothing to recover from the agent's own context.

\paragraph{Diagnosis.}
This is a known trap (carried in the project memory note ``Parallel
agent cap 2--3''): four or more simultaneous agents trip a rate-limit
kill within $\sim3.5$ minutes of sustained fan-out. The kill is
non-graceful --- it does not flush partial work.

\paragraph{Salvage pattern (the standing rule).}
The session survived the kill with \emph{zero loss of landed work}
because of the atomic-commit discipline: any work an agent had already
committed to \texttt{main} was absorbed and intact; only the killed
agent's \emph{uncommitted} in-flight edits were lost, and those were
re-derivable. The standing rules are two:
\begin{enumerate}
  \item \textbf{Cap concurrent agents at $2$--$3$.} Subsequent rounds
        held cap $2$ strictly and saw no further kills.
  \item \textbf{Commit early, commit atomically.} Each agent stages and
        commits its own explicit files as soon as a unit of work is
        verified, so that a mid-task kill costs at most one
        un-committed unit, never a landed one. Salvage is then a matter
        of re-running the killed unit, not reconstructing lost state.
\end{enumerate}

\subsection{The concurrency cap and single-writer discipline on hot files}
\label{app:operational:concurrency}

The $2$--$3$ cap is not only a rate-limit accommodation; it is also a
contention bound on \emph{hot files} --- the small set of
frequently-touched coordination files (domain snapshot, log,
roster). When two agents edit the same hot file concurrently, the second
write can silently wipe the first (a stale-context overwrite). The
operating rule that emerged: hot files have a \emph{single writer per
round}. Parallel agents partition by file --- each agent owns the files
it stages --- and the coordination files are updated by exactly one
agent, sequentially, after the parallel work lands. This is the
file-partition corollary of the @D~d9 worktree contract
(\S\ref{app:operational:d9}).

\subsection{Codegen and diagnosis pitfalls}
\label{app:operational:codegen}

Several findings in the session were \emph{not} physics errors but
artifacts of the compile / numeric / tooling layer that initially
\emph{looked} like physics errors. They are catalogued here because each
cost real investigation time and each became a standing diagnostic
heuristic.

\paragraph{(1) The $4\pi$ / division-vs-multiplication precision trap.}
The first Bosch--Hale port placed $\theta$ in a \emph{multiplicative}
position where the published Eq.~\eqref{eq:boschhale} form has it as a
\emph{denominator}. The result: reactivity came out $\approx4\times$ too
low at \emph{every} temperature. The clean $4\times$ ratio was
seductive --- it looked like a missing $4\pi$ somewhere in a
solid-angle normalization, and an hour was spent hunting the phantom
$4\pi$. The truth was a literal transcription typo: the published form
is a division, not a multiplication. Heuristic: \emph{a suspiciously
clean integer ratio (exactly $4\times$, exactly $2\times$) across all
inputs is far more often a transcription typo than a missing physical
constant.} This trap was \emph{not} logged to the adapter defect catalog
--- it was an authoring typo, not a stdlib defect, and the honesty line
between ``my mistake'' and ``a defect in the code home'' was held.

\paragraph{(2) The constant-fold precision wart and the $\pi$-free rewrite.}
The Wesson cylindrical safety factor written literally as
$q^{*} = 2\pi a^2 B / (\mu_0 R I_p)$ suffered a middle-magnitude
constant-folding precision loss of $|\Delta|\approx5.8\times10^{-6}$:
the engine's constant folder cannot reduce the transcendental $\pi$ to a
finite exact value, so the $2\pi/(4\pi\times10^{-7})$ cancellation is
done in floating point and the two $\pi$'s do not cancel to the last
bit. Encoding the exact identity $2\pi/(\mu_0\!\cdot\!10^{6})=5$ as the
BLUE-MAX atom \verb|q_star_constant| licenses the $\pi$-free form
$q^{*} = 5 a^2 B / (R I_p^{\text{MA}})$, in which no transcendental
survives folding and the recompute is $|\Delta|=0$ on every input. Full
derivation in Appendix~\ref{app:bluemax}. Heuristic: \emph{prefer the
algebraically-reduced ($\pi$-free, integer-coefficient) form of a
closed-form kernel; it is what the constant folder can reproduce
exactly.}

\paragraph{(3) The far-field-truncation diagnosis reversal.}
The GetDP FEM on-axis $B_z$ disagreed with the closed-form anchor by
$\sim1.4\%$. The hypothesis --- pursued for five to six hours --- was a
$2\pi$/\texttt{VolAxiSqu} Jacobian / units bug in the \verb|.pro| deck,
because a $2\pi$-flavored error is exactly what one expects from an
axisymmetric $A_\phi$ formulation. \emph{The diagnosis was wrong.} The
real cause was geometric: the air-region outer boundary was only
$\sim5\times$ the coil radius, leaking return flux. Enlarging
$R_{\text{out}}/H_{\text{out}}$ to $\sim10\times$ collapsed the residual
to $<0.1\%$ with no change to the deck's algebra. The full chase and the
deliberately-recorded false diagnosis live in
Appendix~\ref{app:adapters} (FUSION-DEFECT-003) and
Appendix~\ref{app:magnet}, \S\ref{app:magnet:farfield}. Heuristic:
\emph{a small few-percent residual in a finite-domain FEM is a
far-field-truncation suspect before it is a units suspect; check the
domain size before auditing the Jacobian.}

\paragraph{(4) UTF-8 / emoji: xelatex over pdflatex.}
The tier badges $\tierBlue\tierGreen\tierYellow\tierOrange\tierRed$ are
rendered, in the source, from literal emoji. \texttt{pdflatex} emits a
fatal error on any literal emoji in the source; \texttt{xelatex}
(and \texttt{lualatex}) render them natively as a Unicode glyph. The
standing rule, encoded in the \texttt{Makefile} header: the document
engine is \texttt{xelatex}; \texttt{make LATEX=pdflatex} is permitted
\emph{only} after every literal emoji has been stripped and replaced by a
\verb|\providecommand| tier macro redefined to a non-emoji marker. This
paper takes the macro route (\verb|\tierBlue| \emph{et al.}\ are colored
discs) precisely so the source is engine-portable.

\paragraph{(5) Cross-repo \texttt{gh}-merge PR-number bug.}
When the FreeGS fix (PR~\#1075) was landed, the PR was created from a
demiurge worktree but targeted the hexa-lang repository.
\texttt{gh} resolved the \emph{default} repo as demiurge and locked
another worktree's \texttt{main}. The PR itself landed (merged), but the
worktree-cleanup tail false-failed. The recovery, per the project memory
note ``demiurge worktree PR landing'': do not trust the
\texttt{gh pr create} exit code; verify the merge state with
\texttt{gh pr view}, then perform the worktree / branch cleanup
manually. Standing rule: cross-repo PRs are created with an explicit
\verb|--head| (and an explicit repo where the worktree's default is
ambiguous), and merge state is always confirmed out-of-band.

\paragraph{(6) The \texttt{\textasciitilde{}} shell-expansion pool mis-route.}
A finding from the wider compute-dispatch layer that bit this session:
an unescaped \texttt{\textasciitilde{}} (home-directory tilde) in a
command argument is expanded by the local shell to the \emph{local}
home, then shipped verbatim to a remote pool host where that path does
not exist --- silently mis-routing the job. Standing rule: pool / cloud
command paths are absolute or are quoted to defer expansion to the
remote shell; never let the local shell expand \texttt{\textasciitilde{}}
in an argument destined for a remote host.

\paragraph{(7) The \texttt{ubu-1} hexa segfault.}
The \texttt{hexa verify} host matrix is not uniform: one pool host
(\texttt{ubu-1}) segfaults on the hexa binary and another (\texttt{ubu-2})
is unreliable for it. The standing route for \texttt{hexa verify} is
therefore the \texttt{mini} host (or \texttt{POOL\_DISABLE=1} to force
local execution). This is a host-capability constraint, not a code
defect, and it shapes \emph{where} verification runs even though it does
not change \emph{what} is verified.

\subsection{@D~d9 shared-worktree discipline}
\label{app:operational:d9}

The single most important standing contract for running parallel agents
against a shared code home is the worktree-concurrency rule, @D~d9:

\begin{quote}
\emph{Sequential commit on \texttt{main}; stage and commit one agent at a
time; \texttt{git add <explicit-files>} only --- never} \verb|git add -A|
\emph{or} \verb|git add .|.
\end{quote}

The failure this prevents is an \emph{index leak}: when two concurrent
agents both have un-staged changes and one runs a broad
\verb|git add|, it absorbs the \emph{other} agent's files into its own
commit. The result is a commit that mixes two agents' work, a wiped
expectation for the second agent, and a polluted history. The rule has
three clauses, each load-bearing:
\begin{enumerate}
  \item \textbf{Sequential commit.} Only one agent commits to the shared
        \texttt{main} index at a time; the others wait. This serializes
        the one resource (the index) that cannot be safely shared.
  \item \textbf{One agent at a time.} Staging is not interleaved ---
        agent $A$ fully stages and commits before agent $B$ begins to
        stage --- so the index never holds two agents' changes
        simultaneously.
  \item \textbf{Explicit-path staging only.}
        \verb|git add <explicit-files>| names exactly the files the agent
        owns; the broad forms are forbidden because they cannot
        distinguish ``my work'' from ``the other agent's un-committed
        work.''
\end{enumerate}
In this session, this discipline is what made the rate-limit salvage
(\S\ref{app:operational:ratelimit}) clean: because each agent committed
only its own explicit files, the killed agent could not have leaked
into, or been leaked into by, a survivor. The two contracts compose ---
the $2$--$3$ concurrency cap bounds the contention, and @D~d9 makes the
contention that does occur safe.

\subsection{What was left open}
\label{app:operational:deferred}

For honesty and reproducibility, the session journal also records what
was \emph{not} closed, carried here so the boundary is explicit:
\begin{itemize}
  \item the $\sim475$\,MB ENDF/B-VIII corpus git-LFS ingest (the
        infrastructure landed in PR~\#1090; the data ingest itself is a
        follow-up, HEXA-PORT P4b);
  \item gate (g), the JET D--T measured oracle, is \emph{permanent}
        wet-lab --- it is never closed by a closed-form simulation, by
        construction (the honesty boundary of \S\ref{sec:context});
  \item the p--$^{11}$B reactivity \emph{absolute} value in cm$^3$/s
        (registered only as a reactivity \emph{shape} pending resolution
        of the keV/MeV S-factor unit ambiguity, \S\ref{sec:keystone});
  \item HEXA-PORT P4a (a native Monte Carlo engine), which is a sibling
        domain's follow-up and out of scope for this paper.
\end{itemize}
These are not gaps in the work presented; they are the deliberately
un-closed edges, recorded so that the difference between ``done'' and
``deferred'' is never ambiguous. Source for this entire appendix:
\texttt{companion/session-journal.md}.
