\section{Pull-Request Roll}
\label{app:pr-roll}

This appendix is the transcription of the
\texttt{companion/\allowbreak pr-roll.json} session pull-request roll:
every merged PR that produced this paper, with its \texttt{gh}-measured
diff-stat. The records carry no synthetic data --- \texttt{merged\_at},
\texttt{files\_changed}, and the line counts are read back from
\texttt{gh pr view {-}{-}json}. Fifteen PRs landed: twelve in the
hexa-lang stdlib SSOT (the kernels and adapters), and three in the
demiurge paper/domain home (the domain pointers and absorption
signoffs). They are presented grouped by repository, in dependency
order matching the Reproducibility landing list (\S\ref{sec:repro}).

\subsection{hexa-lang stdlib PRs (the implementation home)}
\label{app:pr-roll:hexalang}

The twelve hexa-lang PRs carry all of the funnel's implementation: the
closed-form kernels, the F-funnel ranker, the three wrap-and-fix
adapter promotions, the closed-form parity anchor, the git-LFS infra,
and the BLUE-MAX audit. Together they add \textbf{3238} lines and remove
\textbf{29} across \textbf{25} file touches (sum of
Table~\ref{tab:app:prroll:hexalang}).

\begin{table}[h]
\centering
\small
\begin{tabular}{rlrrl}
\toprule
\# & merged (UTC) & files & $\pm$ lines & gate \\
\midrule
1012 & 05-25 09:25 & 3 & $+559/-0$  & (a) reactivity + (b) Lawson \\
1020 & 05-25 10:51 & 2 & $+385/-0$  & F-funnel composite \\
1027 & 05-25 11:09 & 4 & $+332/-0$  & (d) MHD stability \\
1032 & 05-25 11:10 & 1 & $+301/-0$  & (c) Grad--Shafranov \\
1042 & 05-25 11:26 & 1 & $+458/-0$  & (e) HTS TF magnet \\
1054 & 05-25 11:44 & 3 & $+241/-1$  & (a) reactivity (aneutronic) \\
1075 & 05-25 12:16 & 1 & $+20/-10$  & (c) Grad--Shafranov \\
1078 & 05-25 12:31 & 1 & $+419/-0$  & (f) TBR neutronics \\
1084 & 05-25 12:39 & 2 & $+38/-0$   & (e) HTS TF magnet \\
1090 & 05-25 12:48 & 2 & $+33/-0$   & infra (HEXA-PORT P4b prep) \\
1095 & 05-25 13:02 & 4 & $+429/-18$ & (e) HTS TF magnet \\
1102 & 05-25 13:18 & 1 & $+23/-0$   & BLUE-MAX audit \\
\bottomrule
\end{tabular}
\caption{The twelve hexa-lang stdlib PRs (dancinlab/hexa-lang), in
dependency order. Diff-stats are \texttt{gh}-measured. The keystone is
\#1012 (the F1/F2 closed-form kernels); the three wrap-and-fix
promotions are \#1075 (FreeGS), \#1078 (OpenMC), and \#1095 (GetDP).}
\label{tab:app:prroll:hexalang}
\end{table}

\paragraph{\#1012 --- F1/F2 keystone closed-form kernels
($+559$, 3 files).}
The Bosch--Hale $\langle\sigma v\rangle(T)$ reactivity (gate~a) and the
Lawson triple-product / $Q_{\text{plasma}}$ closed-form kernels
(gate~b). This is the keystone PR: every other closed-form atom and the
F-funnel ranker depend on it.

\paragraph{\#1020 --- F-funnel composite ranker ($+385$, 2 files).}
The design-point $(a)\!\times\!(b)$ composite ranker that consumes the
\#1012 kernels and emits a top-$K$ JSON of operating points.

\paragraph{\#1027 --- ideal-MHD $\beta$-limit ($+332$, 4 files).}
The gate-(d) Troyon $\beta$-limit, the Wesson cylindrical $q^{*}$, and
the kink criterion as closed-form kernels --- the source of the
load-bearing $q^{*}$-constant identity (Appendix~\ref{app:bluemax}).

\paragraph{\#1032 --- FreeGS Grad--Shafranov wrap ($+301$, 1 file).}
The gate-(c) free-boundary equilibrium wrap, landed initially as an
install-gated honest-skip stub --- the configuration that PR~\#1075
later had to fix.

\paragraph{\#1042 --- HTS REBCO TF-coil field-margin gate
($+458$, 1 file).}
The gate-(e) \texttt{magnet\_gate} adapter, delegating to the RTSC HTS
substrate, with a closed-form $1/R$ field baseline (the pre-bridge
path, before the GetDP FEM bridge of \#1095).

\paragraph{\#1054 --- p--$^{11}$B aneutronic reactivity
($+241/-1$, 3 files).}
The aneutronic fuel axis: the p--$^{11}$B $\langle\sigma v\rangle(T)$
kernel on the Tentori--Belloni 2023 / Nevins--Swain 2000 framework
(reactivity \emph{shape} only --- see \S\ref{sec:keystone}).

\paragraph{\#1075 --- FreeGS wrap-and-fix ($+20/-10$, 1 file).}
The first wrap-and-fix: the \#1032 stub's divergent custom coils were
replaced by the canonical \texttt{TestTokamak} preset with
\texttt{freeBoundaryHagenow} closure, and a stderr redirect stopped the
``No O points'' warning from corrupting the JSON output line. Captures
the P1 parity oracle ($q_{95}\approx10.08$).

\paragraph{\#1078 --- OpenMC TBR real-transport adapter
($+419$, 1 file).}
The second wrap-and-fix: a transport-less \texttt{.py} gate-classifier
stub was superseded by a real \texttt{.hexa} OpenMC adapter that builds
the breeder geometry and runs a 5M-history transport, validating
$\text{TBR}=1.3222\pm0.0004$. Captures the P4 parity oracle.

\paragraph{\#1084 --- closed-form solenoid anchor ($+38$, 2 files).}
The thick-finite-solenoid on-axis $B_z$ closed-form atom
(\verb|solenoid_axis_bz|), registered as the load-bearing parity anchor
for the GetDP FEM bridge --- the HEXA-PORT P2 oracle anchor.

\paragraph{\#1090 --- git-LFS infra ($+33$, 2 files).}
HEXA-PORT P0.5: git-LFS scaffolding (\texttt{.gitattributes} patterns +
README section) for future data-bound corpora, the storage carrier for
the deferred ENDF/B-VIII corpus (Appendix~\ref{app:hexaport}).

\paragraph{\#1095 --- GetDP FEM bridge promote ($+429/-18$, 4 files).}
The third wrap-and-fix: a hexa-native GetDP driver (\texttt{solenoid\_%
solve}) closes the HEXA-PORT P2 parity oracle at $\Delta=-0.064\%$ (the
far-field truncation was raised $5\times\!\rightarrow\!10\times$),
flipping \texttt{magnet\_gate}'s \texttt{field\_path} from the
closed-form baseline to \texttt{"getdp"}.

\paragraph{\#1102 --- BLUE-MAX 8 integer atoms ($+23$, 1 file).}
The BLUE-MAX audit: eight \tierBlue\ \textsc{Supported-Formal} integer
atoms (including \verb|q_star_constant|$=5$) plus eight \tierRed\
matched negative controls (Appendix~\ref{app:bluemax}).

\subsection{demiurge paper/domain PRs (the pointer home)}
\label{app:pr-roll:demiurge}

The three demiurge PRs are the domain-pointer and absorption side: they
register the FUSION and HEXA-PORT domains, sign off the gate-(e) F5
promote and the HEXA-PORT P2 oracle close-out, and absorb the BLUE-MAX
audit. They add \textbf{298} lines and remove \textbf{1} across
\textbf{9} file touches.

\begin{table}[h]
\centering
\small
\begin{tabular}{rlrrl}
\toprule
\# & merged (UTC) & files & $\pm$ lines & role \\
\midrule
176 & 05-25 12:59 & 4 & $+293/-0$ & FUSION + HEXA-PORT domain birth \\
178 & 05-25 13:05 & 3 & $+3/-1$   & F5 promote + P2 oracle close-out \\
179 & 05-25 13:21 & 2 & $+2/-0$   & BLUE-MAX \tierBlue\ audit absorption \\
\bottomrule
\end{tabular}
\caption{The three demiurge PRs (dancinlab/demiurge), the
domain-pointer / absorption side. \#176 births the FUSION and HEXA-PORT
domain snapshots and logs; \#178 signs off the gate-(e) F5 GetDP promote
and the HEXA-PORT P2 oracle; \#179 absorbs the BLUE-MAX audit pointer.}
\label{tab:app:prroll:demiurge}
\end{table}

\paragraph{\#176 --- FUSION + HEXA-PORT domain birth ($+293$, 4 files).}
Adds the FUSION (magnetic-confinement fusion design) and HEXA-PORT
(native-solver port) domain pointers --- snapshots plus append-only
logs --- as sibling domains.

\paragraph{\#178 --- F5 promote + P2 oracle ($+3/-1$, 3 files).}
The gate-(e) F5 GetDP promote signoff and the HEXA-PORT P2 oracle
close-out: the last honest blocker sealed once \#1095 landed.

\paragraph{\#179 --- BLUE-MAX audit absorption ($+2$, 2 files).}
The BLUE-MAX \tierBlue\ audit (\#1102) absorption pointer, closing the
12 libm-class atoms plus the 8 integer attestations.

\subsection{Roll totals}
\label{app:pr-roll:totals}

Across both repositories the session landed \textbf{15} PRs touching
\textbf{34} files, adding \textbf{3536} lines and removing \textbf{30}.
The implementation mass (3238 of the 3536 added lines) is in hexa-lang,
the canonical code home (@D~d3); the demiurge side is pointer and
absorption metadata only. Three of the twelve hexa-lang PRs (\#1075,
\#1078, \#1095) are wrap-and-fix promotions --- each one surfaced a real
adapter defect on first execution (Appendix~\ref{app:adapters}).
