\section{Reproducibility Runbook}
\label{app:repro}

This appendix is the complete reproducibility runbook. It gives, for
every atom in the verification ledger
(Appendix~\ref{app:verify-ledger}), the exact \texttt{hexa verify
{-}{-}expr <fn> <args> <expected>} CLI line that recomputes it, so a
reader can re-run any single claim in one command. It then pins the
build-from-canonical-root procedure, the external-backend environment
specifications, the canonical data paths, and the \texttt{make} /
\texttt{/paper lint} targets. The data source is the
\texttt{verify-ledger.json} record, the \texttt{companion/README.md},
and the paper Makefile.

\subsection{Build from the canonical root}
\label{app:repro:build}

Every kernel cited in this paper lives under hexa-lang \texttt{stdlib/}
at a single canonical home (@D~d3) and is reachable from the top-level
\texttt{hexa verify} CLI. Two operational notes carry over from the
sibling domains' build experience:
\begin{itemize}\itemsep0.2em
\item \textbf{Build from the canonical hexa-lang root.} The
\texttt{hexa verify} binary must be built from \texttt{\textasciitilde
/core/hexa-lang}; a \texttt{/tmp} worktree skips the \texttt{use}-path
expansion and the atom recompute will not resolve. hexa-lang is
compiled (\texttt{.hexa} $\rightarrow$ C $\rightarrow$ binary), linked
with \texttt{-lm -ldl}.
\item \textbf{Verify host pin.} Atom recompute is routed to the mini
host (\texttt{POOL\_DISABLE=1}); the ubu pool hosts are not the verify
host for this runbook.
\end{itemize}

\subsection{Numerical atoms --- gate (a)/(b)/(d) closed-form}
\label{app:repro:numerical}

The closed-form numerical atoms (\tierGreen, \textsc{Supported-Numerical})
recompute libm-class formulae to a discriminating epsilon. Each line is
\texttt{verify {-}{-}expr <fn> <args> <expected>}; all return
\textsc{Supported-Numerical} at the $|\Delta|$ recorded in the ledger.

\begin{lstlisting}
# gate (a) Bosch-Hale reactivity (PR #1012)
verify --expr sigma_v_dt_e18 10.0  113.617     # |Delta|=0
verify --expr sigma_v_dt_e18 20.0  433.02      # |Delta|=0
verify --expr sigma_v_d3he_e21 50.0  21.4      # |Delta|=0
verify --expr sigma_v_dd_p 15.0  5.781e-18     # |Delta|=0
verify --expr sigma_v_dd_n 15.0  5.435e-18     # |Delta|=0

# gate (b) Lawson / 0-D gain (PR #1012)
verify --expr triple_product 1.0e20 15.0 3.0  4.5e21   # |Delta|=0
verify --expr q_plasma 50.0 10.0  5.0                  # |Delta|=0
verify --expr ignition_margin 50.0 100.0 3.0 0.2  0.3  # |Delta|=0

# F-funnel composite ranker (PR #1020)
verify --expr f_reactivity_coverage 4.330e-16 8.94e-16  0.4843  # |Delta|=0
verify --expr f_margin_coverage 0.3 0.3  1.0                    # |Delta|=0
verify --expr f_funnel_composite 0.4843 1.0  0.4843             # |Delta|=0

# gate (d) ideal-MHD beta-limit (PR #1027)
verify --expr troyon_beta_max 2.8 15.0 5.3 1.6  0.0494   # |Delta|=0
verify --expr q_star_cyl 2.0 5.3 6.2 15.0e6  2.667       # |Delta|=0

# gate (a) aneutronic p-11B (PR #1054)
verify --expr pb11_gamow_factor 148.0  0.0              # |Delta|=0
verify --expr pb11_sfactor_resonance 148.0  18400.0     # |Delta|=0
verify --expr pb11_reactivity_moment_e6 123.0  0.6432   # |Delta|=0
\end{lstlisting}

\subsection{Numerical atoms --- gate (c)/(e)/(f) wrap-run oracles}
\label{app:repro:oracles}

These three atoms are the captured parity oracles
(Appendix~\ref{app:hexaport}, Table~\ref{tab:app:hexaport:oracles}).
Their arguments are the run configuration rather than scalar inputs,
and their recompute requires the external backend installed (the
environment specs of \S\ref{app:repro:env}); without it the adapter
honest-skips rather than fabricating a value.

\begin{lstlisting}
# gate (e) closed-form solenoid anchor (PR #1084) -- no backend needed
verify --expr solenoid_axis_bz 2.0e6 0.50 0.70 1.20  1.48265   # |Delta|=0

# gate (e) GetDP FEM B_peak, Delta = -0.064% vs anchor (PR #1095)
verify --expr magnet_gate_b_peak "iter-class" "getdp"  1.48168  # |Delta|=0.001

# gate (c) FreeGS free-boundary equilibrium q95 (PR #1075)
verify --expr freegs_equilibrium_q95 "TestTokamak" "freeBoundaryHagenow" 10.08
#   |Delta|=0.005 ; Ip=200 kA, q0=1.355, q95~10.08, kappa=1.358

# gate (f) OpenMC TBR, 5M histories (PR #1078)
verify --expr openmc_tbr_pb157li "Pb-15.7Li" "Li6-90at%" "5M-histories" 1.3222
#   |Delta|=0.0004 ; W armor / EUROFER FW / SS316 shield ; TBR > 1
\end{lstlisting}

\subsection{Formal atoms --- BLUE-MAX algebraic roots (PR~\#1102)}
\label{app:repro:formal}

The eight \tierBlue\ \textsc{Supported-Formal} integer atoms are
$0$-argument recomputes; each returns the exact integer at
$|\Delta|=0$ by construction (Appendix~\ref{app:bluemax}).

\begin{lstlisting}
verify --expr bosch_hale_gamow_factor()              = 3
verify --expr sigma_v_dt_e18_scale_exponent()        = 18
verify --expr pb11_gamow_eg_kev()                    = 22589
verify --expr pb11_resonance_kev_doubled()           = 296
verify --expr pb11_quadrature_nodes()                = 600
verify --expr pb11_reactivity_moment_e6_scale_exp()  = 6
verify --expr solenoid_axis_bz_half_factor()         = 2
verify --expr q_star_constant()                      = 5
\end{lstlisting}

\subsection{Negative-control atoms --- BLUE-MAX (PR~\#1102)}
\label{app:repro:negative}

The eight matched negative controls assert the true integer $\pm1$ and
\emph{must} return \tierRed\ \textsc{Falsified} at $|\Delta|=1$. A
runbook that cannot reproduce these falsifications has not reproduced
the discriminating behaviour of the rubric.

\begin{lstlisting}
verify --expr bosch_hale_gamow_factor_NEG()              = 4      # FALSIFIED (true 3)
verify --expr sigma_v_dt_e18_scale_exponent_NEG()        = 17     # FALSIFIED (true 18)
verify --expr pb11_gamow_eg_kev_NEG()                    = 22590  # FALSIFIED (true 22589)
verify --expr pb11_resonance_kev_doubled_NEG()           = 297    # FALSIFIED (true 296)
verify --expr pb11_quadrature_nodes_NEG()                = 601    # FALSIFIED (true 600)
verify --expr pb11_reactivity_moment_e6_scale_exp_NEG()  = 7      # FALSIFIED (true 6)
verify --expr solenoid_axis_bz_half_factor_NEG()         = 3      # FALSIFIED (true 2)
verify --expr q_star_constant_NEG()                      = 6      # FALSIFIED (true 5)
\end{lstlisting}

This is the full 36-atom ledger: 20 \tierGreen\
\textsc{Supported-Numerical} + 8 \tierBlue\ \textsc{Supported-Formal} +
8 \tierRed\ \textsc{Falsified} controls. Each line above corresponds
1:1 to a \texttt{verify-ledger.json} entry, queryable with a one-line
\texttt{jq} filter, e.g.\
\verb#jq '.[] | select(.fn=="q_star_constant")'#.

\subsection{External-backend environment specifications}
\label{app:repro:env}

The three wrap-run oracle atoms (\S\ref{app:repro:oracles}) require an
external backend. The pinned environments are:

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\toprule
backend & version & key dependency / data \\
\midrule
FreeGS & 0.8.2 & numpy $<\!2.0$ \\
OpenMC & 0.15.3 & ENDF/B-VIII.0 (via micromamba) \\
GetDP  & 3.5.0 / 4.0.0 & gmsh (mesh generation) \\
\bottomrule
\end{tabular}
\caption{External-backend environment specifications for the three
wrap-run oracle atoms. FreeGS pins numpy below 2.0; OpenMC needs the
ENDF/B-VIII.0 evaluated-data library installed via micromamba; GetDP
needs gmsh for axisymmetric mesh generation.}
\label{tab:app:repro:env}
\end{table}

\begin{itemize}\itemsep0.2em
\item \textbf{FreeGS 0.8.2} (\verb|freegs_equilibrium_q95|): a Python
package; the numpy~$<\!2.0$ constraint avoids the deprecated-API breaks
in the FreeGS 0.8.2 release. Install into a dedicated environment so the
numpy pin does not collide with other tools.
\item \textbf{OpenMC 0.15.3 + ENDF/B-VIII.0}
(\verb|openmc_tbr_pb157li|): installed via micromamba; the ENDF/B-VIII.0
cross-section library is the data-bound dependency
(Appendix~\ref{app:hexaport}, Tier~C). The adapter builds the geometry
in-process but reads cross sections from the library, so
\verb|OPENMC_CROSS_SECTIONS| must point at the installed
\texttt{cross\_sections.xml}.
\item \textbf{GetDP 3.5.0 / 4.0.0 + gmsh} (\verb|magnet_gate_b_peak|):
either GetDP point release works; gmsh generates the axisymmetric mesh
the \texttt{.pro}/\texttt{.geo} decks consume. The far-field truncation
must be at $\sim$10$\times$ the coil radius (not $5\times$) for the
$\Delta=-0.064\%$ parity --- see the GetDP defect entry
(Appendix~\ref{app:adapters}).
\end{itemize}

Without the relevant backend, the wrap adapter reports an install-gated
honest-skip; it never fabricates a value. The closed-form anchor
(\verb|solenoid_axis_bz|) and all formal/negative atoms need no backend
and recompute on the bare hexa-lang build.

\subsection{Canonical data paths}
\label{app:repro:data}

\begin{itemize}\itemsep0.2em
\item \textbf{Kernels.} hexa-lang \texttt{stdlib/} (the single
canonical home, @D~d3); the FUSION kernels under
\texttt{stdlib/fusion/}, the magnet bridge under \texttt{stdlib/rtsc/}.
\item \textbf{GetDP decks.} The \texttt{.pro} / \texttt{.geo} decks
ship under \texttt{stdlib/rtsc/decks/}; they are kilobyte text and are
\emph{not} git-LFS gated.
\item \textbf{Deferred ENDF/B-VIII corpus.} The future hexa-native
cross-section corpus lands under \texttt{data/nuclear/endf\_b8\_0/} on
hexa-lang git-LFS (infra in PR~\#1090; the $\sim$475\,MB ingest is a
separate HEXA-PORT P4b follow-up --- Appendix~\ref{app:hexaport}).
\item \textbf{Data-record axis.} The paper's \texttt{companion/}
directory carries \texttt{verify-ledger.json} (the 36-atom ledger),
\texttt{pr-roll.json} (the 15 PRs, Appendix~\ref{app:pr-roll}),
\texttt{adapter-defect-catalog.json} (the three wrap-and-fix defects,
Appendix~\ref{app:adapters}), and \texttt{session-journal.md}.
\end{itemize}

\subsection{Build and lint targets}
\label{app:repro:make}

The paper builds with the Makefile in the paper root:

\begin{lstlisting}
make            # main.pdf : xelatex x3 + bibtex (cross-ref resolution)
make figures    # regenerate figures/*.pdf from figures/_scripts/*.{py,tex}
make pages      # report page count via pdfinfo
make wordcount  # word count (texcount, falls back to wc -w)
make lint       # /paper lint . (commons @D g51: >=10 pages + >=1 figure)
make arxiv-tar  # out/<slug>-arxiv.tar.gz : arxiv-ready source bundle
make clean      # remove intermediates (PDF preserved)
\end{lstlisting}

Two build notes carry the engine constraint. First, the engine is
\textbf{xelatex} (\texttt{LATEX = xelatex {-}interaction=nonstopmode
{-}halt-on-error}): the tier badges are rendered as TikZ colored discs
(\verb|\tierBlue|\,\dots\verb|\tierRed|) so the build is engine-agnostic,
but a \texttt{make LATEX=pdflatex} override requires first stripping any
literal emoji from the source. Second, \texttt{make} runs
\texttt{xelatex} three times around a single \texttt{bibtex} pass to
resolve the cross-references and the \texttt{natbib} citations; the
\texttt{main.bbl} is committed so a single-pass rebuild still resolves
citations if \texttt{bibtex} is unavailable. The \texttt{make lint}
target enforces the commons @D~g51 floor ($\geq$10 pages and at least
one figure).
