% FILLED (fill batch): J — inherited RTSC magnet toolchain.
%   folds the RTSC inheritance provenance — solenoid_axisym .geo/.pro (GetDP)
%   + the Wheeler on-axis closed-form anchor + the 5-axis RtscVerifyRecord
%   schema extension (device=ioffe_pritchard_trap), from RTSC.md §4.2 and
%   ~/core/hexa-lang/stdlib/rtsc/ + stdlib/material/magnet/.
\section{Inherited RTSC Magnet Toolchain}
\label{app:rtsc}

This appendix documents the cross-domain magnet inheritance that powers
process \textcircled{\scriptsize 6} (Appendix~\ref{app:confinement}). We trace
each reused asset to its RTSC origin, argue why the same magnetostatic spine
serves a field \emph{well} (this antihydrogen trap) and a field \emph{peak}
(an RTSC high-field solenoid) with only a coil-geometry change, and give the
deck-derivation runbook. Provenance: \texttt{RTSC.md} \S4.2, hexa-lang
\texttt{stdlib/rtsc/} and \texttt{stdlib/material/magnet/}.

% -------------------------------------------------------------------------
\subsection{Asset-by-asset provenance}
\label{app:rtsc-assets}

The confinement stage adds \emph{no new solver and no new closed form}. Every
magnetostatic asset it uses already existed in the RTSC magnet domain and is
reused unmodified; Table~\ref{tab:rtsc-assets} traces each to its canonical
stdlib home.

\begin{table}[h]
\centering
\small
\setlength{\tabcolsep}{5pt}
\begin{tabular}{@{}>{\raggedright\arraybackslash}m{4.0cm} >{\raggedright\arraybackslash}m{4.3cm} >{\raggedright\arraybackslash}m{3.6cm}@{}}
\toprule
\textbf{RTSC asset} & \textbf{canonical home} & \textbf{ANTIMATTER reuse} \\
\midrule
Wheeler on-axis $B$ closed form (current-loop primitive at $r=0$) &
\texttt{stdlib/material/magnet/} &
the $B_{\mathrm{loop}}(\zeta)$ anchor of Eq.~(F.\ref{eq:conf-loop}),
superposed into a magnetic well \\[3pt]
\texttt{solenoid\_axisym.geo} (OCC + BooleanFragments parametric mesh) &
\texttt{stdlib/rtsc/templates/} &
re-parametrized into a two-coil mirror-pair geometry \\[3pt]
\texttt{solenoid\_axisym.pro} (GetDP Form1P A--$\phi$, axis Dirichlet) &
\texttt{stdlib/rtsc/templates/} &
the same axisymmetric magnetostatic formulation, mirror layout \\[3pt]
\texttt{getdp\_hts.py} 5-axis driver (closed-form V1 $\times$ GetDP FEM V2 + cross-check) &
\texttt{stdlib/rtsc/} &
the verify-once / reuse driver (FEM re-solve deferred, App.~\ref{app:confinement}) \\[3pt]
\texttt{RtscVerifyRecord} 5-axis schema &
\texttt{cockpit/.../Models/} &
extended with \texttt{device = ioffe\_pritchard\_trap} \\
\bottomrule
\end{tabular}
\caption{Direct inheritance of the RTSC magnet toolchain at process
\textcircled{\scriptsize 6}. The Wheeler on-axis closed form and the GetDP
axisymmetric A--$\phi$ deck (\texttt{RTSC.md} \S4.2) are reused unmodified;
only the coil layout (two mirror coils instead of one solenoid) and the
record's \texttt{device} field change.}
\label{tab:rtsc-assets}
\end{table}

% -------------------------------------------------------------------------
\subsection{The Wheeler closed form and its RTSC cross-check}
\label{app:rtsc-wheeler}

The single load-bearing primitive is the on-axis current-loop field
$B_{\mathrm{loop}}(\zeta)=\mu_0 I a^2/[2(a^2+\zeta^2)^{3/2}]$
(Eq.~(F.\ref{eq:conf-loop})). In the RTSC domain this same Wheeler on-axis
form was FEM-cross-checked against a GetDP axisymmetric A--$\phi$ solve of a
representative solenoid (200\,mm length, $r\in[30,55]$\,mm, 120 turns,
$I=100$\,A): the on-axis centre field agreed to
\begin{equation}
B_{\mathrm{center}}^{\mathrm{Wheeler}} = 0.06939\ \mathrm{T},
\quad
B_{\mathrm{center}}^{\mathrm{FEM}} = 0.06842\ \mathrm{T},
\quad
\Delta = -1.40\%
\label{eq:rtsc-xcheck}
\end{equation}
(\texttt{RTSC.md} \S4.2.1, GetDP 3.5.0 cross-check record; the current toolchain
binary is GetDP 4.0.0 ARM \citep{getdp}). The $-1.40\%$ on-axis $B$ agreement
is exactly the quantity ANTIMATTER inherits, because the Ioffe--Pritchard
trap depth depends only on \emph{on-axis} $B_{\min}$ and $B_{\mathrm{coil}}$.
We note honestly that RTSC's \emph{inductance} cross-check showed a larger
$-21\%$ Wheeler-vs-FEM gap, attributed to the thin-coil approximation in the
Wheeler inductance formula at a thick aspect ratio ($b/a=1.83$) --- but
inductance is not used by the trap-depth calculation, so that caveat does
\emph{not} propagate to process \textcircled{\scriptsize 6}.

% -------------------------------------------------------------------------
\subsection{Peak versus well: one spine, two targets}
\label{app:rtsc-peak-well}

A solenoid \emph{concentrates} field: a single coil (or close-packed stack)
makes $|B|$ maximal at the centre. An Ioffe--Pritchard trap \emph{minimizes}
it: two coils pulled apart ($s>a$) make $|B|$ minimal at the centre and rising
outward. These are the \emph{same superposition of the same Wheeler primitive}
with a different coil separation --- Helmholtz-like ($s<a$, a peak) versus
mirror ($s>a$, a well). The magnetostatic kernel verified once in RTSC (the
on-axis $B$ agreement of Eq.~\eqref{eq:rtsc-xcheck}) is therefore re-verified
here against a \emph{new physical observable} --- a neutral-atom trap depth in
kelvin (Eq.~(F.\ref{eq:conf-depth})) --- with no new code. This is the
verify-once / reuse-twice payoff of the hexa-lang stdlib SSOT discipline: a
primitive is verified at its canonical home and then serves any geometry that
superposes it.

% -------------------------------------------------------------------------
\subsection{Deck-derivation runbook}
\label{app:rtsc-runbook}

To reproduce the optional GetDP FEM re-solve of the mirror geometry (the
deferred cross-check of App.~\ref{app:confinement}):
\begin{enumerate}\itemsep0.2em
\item Copy \texttt{stdlib/rtsc/templates/solenoid\_axisym.geo} and set two
  coil bands at $z=\pm s/2$ (instead of one centred band).
\item Reuse \texttt{solenoid\_axisym.pro} verbatim (Form1P A--$\phi$,
  axisymmetric, axis Dirichlet) --- the formulation is geometry-independent.
\item Solve with GetDP 4.0.0 ARM (\texttt{\$GETDP\_BIN}); the on-axis
  $B_{\min}$/$B_{\mathrm{coil}}$ should reproduce the closed-form
  Table~(F.\ref{tab:conf-verify}) values within the $\sim$1.4\% on-axis FEM
  agreement of Eq.~\eqref{eq:rtsc-xcheck}.
\item Emit an \texttt{RtscVerifyRecord} with
  \texttt{device = ioffe\_pritchard\_trap}.
\end{enumerate}

% -------------------------------------------------------------------------
\subsection{Honest scope}
\label{app:rtsc-scope}

The inheritance is of the \emph{on-axis} closed form and its RTSC cross-check
($-1.40\%$ on $B_{\mathrm{center}}$). The FEM re-solve of the mirror geometry
is documented but \emph{not run} here (deferred, App.~\ref{app:confinement});
the inductance-formula caveat ($-21\%$, thin-coil at thick aspect) is recorded
for honesty but is irrelevant to the trap-depth chain. The RTSC records
themselves carry \texttt{absorbed=false} (HTS critical-state not modelled in
the linear A--$\phi$ solve); ANTIMATTER inherits that honesty stance unchanged.
