\section{pylhe LHE Event-Stats --- Full Data Record}
\label{app:lhe}

This appendix is the full data record for the LHE event-analysis stage
(pipeline stage \textcircled{\scriptsize 7}, the \emph{analyze} verb,
\S\ref{sec:fodo}). It states the synthetic
$e^+e^-\!\to Z\!\to\mu^+\mu^-$ sample at $\sqrt{s}=\SI{91.1876}{GeV}$
(100 events), the per-event kinematic statistics the pylhe\,1.0.4 cell
computes, and the \emph{permanent} \texttt{absorbed=false} /
\textsc{Gate-Open} verdict: the events are synthetic (not tuned Monte
Carlo) and there is no detector simulation, so the cell is a
demonstrated parser surface, not a physics claim.

% -------------------------------------------------------------------------
\subsection{Sample and process}
\label{app:lhe-sample}

The cell (\verb|stdlib/cern/lhe_stats.py|, pylhe\,1.0.4, Python\,3.14.4)
reads a synthetic Les Houches Event v3.0 file and computes per-event
kinematic statistics. Table~\ref{tab:app-lhe-proc} is the process
definition, reproduced from the export record
(\verb|cern_lhe_20260519T172736Z.json|).

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}l l@{}}
\toprule
field & value \\
\midrule
channel              & $e^+e^-\to Z\to\mu^+\mu^-$ \\
$\sqrt{s}$           & \SI{91.1876}{GeV} (on the $Z$ pole) \\
events               & 100 (requested $=$ produced) \\
LHE version          & 3.0 \\
RNG seed             & 20260520 (deterministic) \\
parser               & pylhe\,1.0.4 (BSD-3) \\
artifact size        & \SI{75211}{bytes} \\
\bottomrule
\end{tabular}
\caption{LHE sample definition. The events are synthetic on-shell $Z$
decays with deterministic decay angles, written to standard LHE v3.0 and
read back through pylhe.}
\label{tab:app-lhe-proc}
\end{table}

% -------------------------------------------------------------------------
\subsection{Per-event kinematic summary}
\label{app:lhe-stats}

Table~\ref{tab:app-lhe-meas} is the complete measurement block. The
final-state energy sum is the strongest internal check:
$100 \times \SI{91.1876}{GeV} = \SI{9118.76}{GeV}$, i.e.\ per-event
4-momentum conservation holds exactly across the round-trip. The
multiplicity is uniformly 5 (the $Z$ plus the four leptons in the LHE
particle list), and the PID counts confirm the channel.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}l r@{}}
\toprule
quantity & value \\
\midrule
events $n$                          & 100 \\
final-state particles              & 200 \\
total final-state energy           & \SI{9118.76}{GeV} \\
total final-state $|p_z|$          & \SI{4594.7668}{GeV} \\
multiplicity (min / mean / max)    & 5 / 5 / 5 \\
PID counts ($e^-$/$e^+$)           & 100 / 100 \\
PID counts ($\mu^-$/$\mu^+$)       & 100 / 100 \\
PID counts ($Z$, pid 23)           & 100 \\
\bottomrule
\end{tabular}
\caption{pylhe per-event kinematic summary (verbatim from the export
\texttt{measurements} block). The energy-sum identity
$9118.76 = 100\times 91.1876$ is the parser-round-trip integrity check.}
\label{tab:app-lhe-meas}
\end{table}

% -------------------------------------------------------------------------
\subsection{Permanent GATE\_OPEN --- the honest boundary}
\label{app:lhe-honesty}

\noindent\fbox{\parbox{0.97\linewidth}{\small\textbf{Permanent
\texttt{absorbed=false} / \textsc{Gate-Open}.} The numbers above are
\emph{real} --- they are the algorithm output of pylhe parsing a standard
LHE v3.0 file. But the verdict is held at \textsc{Gate-Open} permanently,
for three structural reasons. (i) \textbf{Synthetic input.} The events
are synthetic kinematics with deterministic decay angles, \emph{not} the
output of a tuned generator (MadGraph / POWHEG / Sherpa); the
distributions (rapidity, $p_T$, angular) are not physical. (ii)
\textbf{No detector simulation.} There is no Delphes/Geant4 smearing,
efficiency, acceptance, or trigger model, so the record cannot be
compared to ATLAS/CMS object selection. (iii) \textbf{Scope.} The cell is
a demonstrated \emph{parser surface} --- a single-point LHE round-trip
--- not a physics prediction. The BSD-3 pylhe + LHE standard are
absorbable as tooling, but the \emph{predictive} content needs a tuned
generator $+$ detector sim $+$ real LHC data, all out of the clean-room
scope. This gate does not graduate.}}
