\section{7-Verb Stub Status --- Honest-Skip Record}
\label{app:stubs}

This appendix records the status of the four \emph{honest-stub} verbs
(specify, structure, design, handoff; Table~\ref{tab:verbs}). For each,
the typed cockpit record (\verb|Cern*Record.swift|) and the
\verb|cern.demi| STUB cell slot exist, but the stdlib substrate
(\verb|stdlib/cern/{specify,structure,design,handoff}.py|) is unwritten,
so the manifest-driven generic dispatch
(\verb|CellrunDispatch.run("cern", verb)|, @D d4 --- zero new producer
classes) returns \texttt{rc=2} (honest-skip) rather than a fabricated
verdict. The \verb|accel_mechanism| field is carried in every record to
make the verb $\times$ mechanism axis split explicit at the data level.

% -------------------------------------------------------------------------
\subsection{Per-verb record schema}
\label{app:stubs-schema}

Each stub carries the common record envelope (\verb|domain|, \verb|verb|,
\verb|kind|, \verb|stamp|, \verb|producer|, \verb|accelMechanism|,
\verb|measurementGate|, \verb|absorbed|, \verb|scopeCaveats|,
\verb|citations|, \verb|skippedReason|, \verb|headline|) plus a small set
of verb-specific headline slots. Table~\ref{tab:app-stubs} lists the
distinguishing slots and the cell's stdlib substrate target.

\begin{table}[h]
\centering
\small
\setlength{\tabcolsep}{5pt}
\begin{tabular}{@{}>{\raggedright\arraybackslash}m{1.6cm} >{\raggedright\arraybackslash}m{4.3cm} >{\raggedright\arraybackslash}m{4.3cm} c@{}}
\toprule
verb & record (Swift) & headline slots & substrate \\
\midrule
specify   & \texttt{CernSpecifyRecord}   & \texttt{targetFieldT}, \texttt{beamEnergyGeV}, \texttt{luminosityCm2S}, \texttt{machineClass} & unwritten \\
structure & \texttt{CernStructureRecord} & \texttt{latticeType}, \texttt{cellCount}, \texttt{circumferenceM} & unwritten \\
design    & \texttt{CernDesignRecord}    & \texttt{deckBackend}, \texttt{deckPath}, \texttt{elementCount} & unwritten \\
handoff   & \texttt{CernHandoffRecord}   & \texttt{deliverables[]}, \texttt{slot}, \texttt{present}, \texttt{bundlePath} & unwritten \\
\bottomrule
\end{tabular}
\caption{The four honest-stub verb records. The CDR target slots
(specify), lattice topology slots (structure), MAD-X deck slots (design),
and TDR deliverable schema (handoff) are typed and decode-ready; the
stdlib \texttt{.py} substrate that would populate them is not yet
authored.}
\label{tab:app-stubs}
\end{table}

% -------------------------------------------------------------------------
\subsection{The \texttt{cern.demi} STUB cells}
\label{app:stubs-cell}

Each stub is a manifest cell in \verb|domains/cern.demi| with a
\verb|script| target, a \verb|record_kind|, and an \verb|output_dir|, but
no implemented backend. The four UNWIRED cells are:

\begin{lstlisting}
[cell.specify]    script = stdlib/cern/specify.py    record_kind = cern_specify_record
[cell.structure]  script = stdlib/cern/structure.py  record_kind = cern_structure_record
[cell.design]     script = stdlib/cern/design.py     record_kind = cern_design_record
[cell.handoff]    script = stdlib/cern/handoff.py     record_kind = cern_handoff_record
\end{lstlisting}

Each carries an honest \verb|scope_caveat| naming the intended producer,
e.g.\ for specify: \emph{``physics case / Conceptual Design Report (CDR)
template producer not authored''}; for structure: \emph{``ring/lattice
topology producer (FODO chain or DBA cell counter) not authored''}; for
design: \emph{``MAD-X lattice deck template producer not authored''}; for
handoff: \emph{``Technical Design Report (TDR) bundle producer not
authored''}. (The design cell notes that the \emph{synthesize} verb's
\texttt{xsuite\_optics.py} acts as a de-facto stand-in for the
matched-optics step; design proper would emit the SEQUENCE deck
\emph{before} matching.)

% -------------------------------------------------------------------------
\subsection{rc=2 honest-skip --- the design note}
\label{app:stubs-rc2}

When the generic dispatch runs a stub cell, it finds the
\verb|script| target missing and returns \texttt{rc=2} (honest-skip):

\begin{lstlisting}
$ cellrun cern specify
[cellrun] cern.specify -> stdlib/cern/specify.py
[cellrun] substrate not found (UNWIRED cell) -> rc=2 (honest-skip)
          measurement_gate = GATE_OPEN, absorbed = false, no record emitted
\end{lstlisting}

All four stub cells return the same honest-skip across the generic
dispatch:

\begin{lstlisting}
$ for v in specify structure design handoff; do cellrun cern $v; done
[cellrun] cern.specify   -> stdlib/cern/specify.py    UNWIRED -> rc=2
[cellrun] cern.structure -> stdlib/cern/structure.py  UNWIRED -> rc=2
[cellrun] cern.design    -> stdlib/cern/design.py     UNWIRED -> rc=2
[cellrun] cern.handoff   -> stdlib/cern/handoff.py    UNWIRED -> rc=2
# all four: measurement_gate=GATE_OPEN, absorbed=false, no record emitted
\end{lstlisting}

\noindent\fbox{\parbox{0.97\linewidth}{\small\textbf{Honest stub
$\ne$ fabricated verdict.} This is the structural payoff of the single
generic dispatch (@D d4). Because every verb of every domain traverses
\emph{one} path --- \texttt{CellrunDispatch.run("cern", verb)} with zero
per-verb producer classes and zero name-branching --- a stub can be
\emph{honest}: the generic path runs the cell, finds no substrate, and
returns \texttt{rc=2} rather than synthesizing a plausible-looking number.
The typed record and the manifest slot are real (decode-ready,
\texttt{accel\_mechanism}-carrying); only the physics substrate is
deferred. A fabricated \tierGreen\ on an unwritten cell would be a g3
violation; \texttt{rc=2} is the correct, auditable state.}}

% -------------------------------------------------------------------------
\subsection{The intended handoff (TDR) deliverable schema}
\label{app:stubs-tdr}

The handoff record is the most structured of the four stubs, since the
Technical Design Report (TDR) it would emit has a known shape. The
\verb|CernHandoffRecord| carries a \verb|deliverables[]| array of
\verb|CernHandoffDeliverable| entries, each a \verb|slot| with a
\verb|present| flag, plus a \verb|bundlePath| to the assembled archive.
The natural deliverable set (from \texttt{CERN.md}, by analogy to the
firmware CycloneDX bundle pattern) is in Table~\ref{tab:app-tdr}.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}l l@{}}
\toprule
slot & content (when authored) \\
\midrule
\texttt{lattice\_deck}      & MAD-X SEQUENCE deck (from the design verb) \\
\texttt{shielding\_dossier} & Geant4 radiation / shielding budget \\
\texttt{da\_scan}           & dynamic-aperture scan results \\
\texttt{rf\_spec}           & RF cavity voltage / frequency spec \\
\texttt{radiation\_budget}  & integrated dose / activation estimate \\
\bottomrule
\end{tabular}
\caption{The TDR deliverable schema the \texttt{handoff} cell would
populate. Today all slots are \texttt{present=false} (substrate
unwritten); the typed shape is decode-ready so a future producer fills it
without a record-schema change.}
\label{tab:app-tdr}
\end{table}
