\section{Downstream / Honest-Scope Catalog}
\label{app:downstream}

This appendix is the explicit catalog of items that are out of the
demiurge clean-room scope --- \emph{not} domain incompleteness, but
external-dependency hand-offs past the tabletop completion criterion
(\S\ref{sec:limits}). The clean-room scope is orthogonal to wet-lab /
GPU / sourced-deck dependency; this catalog names each boundary, its
exact external dependency, and the appropriate hand-off path. This is the
signature honesty axis of the CERN domain: the scope boundary is a
\emph{deliverable}, not an apology.

% -------------------------------------------------------------------------
\subsection{The three downstream items}
\label{app:downstream-table}

\begin{table}[h]
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}>{\raggedright\arraybackslash}m{3.2cm} >{\raggedright\arraybackslash}m{5.1cm} >{\raggedright\arraybackslash}m{4.6cm}@{}}
\toprule
\textbf{item} & \textbf{external dependency} & \textbf{appropriate hand-off path} \\
\midrule
Geant4-MC stopping-power solution (Stage 4)
  & Geant4 source build $+$ \texttt{geant4\_pybind} trampoline ownership
    (pybind11 segfault surfaced in an earlier round); the four corrections
    (shell, density, straggling, nuclear)
  & wet-lab-equivalent: GPU pod cycle or \texttt{/micro-exp} sweep, against
    the Bethe--Bloch Stage-3 kernel (Appendix~\ref{app:bethe}) \\[4pt]
measured-ring optics parity
  & sourced LHC / FCC-ee / SPS optics deck (licensing $\cdot$ clean-room
    risk) $+$ a measured tune
  & external-data ingest as a \emph{separate} domain, on the facility's
    official release (extends Appendix~\ref{app:fodo}) \\[4pt]
2-D nonlinear blowout PIC
  & 2-D / 3-D FBPIC / WarpX GPU run with an $n_p$ / drive-intensity sweep
  & \texttt{/micro-exp} GPU sweep, a design-grade separate cycle, beyond
    the 1-D linear parity (Appendix~\ref{app:pic}) \\
\bottomrule
\end{tabular}
\caption{The \texttt{CERN.md \#\# downstream} catalog. Each row is a
\emph{scope boundary}, never an impossibility: the closed-form / parity
work inside the clean room is complete to its tabletop terminus, and each
downstream item has a named external dependency and a concrete hand-off
path.}
\label{tab:app-downstream}
\end{table}

% -------------------------------------------------------------------------
\subsection{Why each is out of scope (not impossible)}
\label{app:downstream-why}

\paragraph{Geant4-MC Stage-4 stopping power.} The Bethe--Bloch kernel
(Appendix~\ref{app:bethe}) is Stage-3 \tierGreen: a
hexa$\leftrightarrow$Python parity on the PDG closed form. Flipping
\texttt{absorbed=true} requires a full Geant4-MC \texttt{G4hIonisation}
parity that adds the four corrections the closed form omits. The blocker
is \emph{toolchain ownership}, not physics: the Geant4 source build plus
the \texttt{geant4\_pybind} trampoline surfaced a pybind11 segfault in an
earlier round. The breakthrough path is concrete --- run Geant4 on a
dedicated GPU pod or via a \texttt{/micro-exp} sweep where the segfault
can be owned and debugged in isolation. The \texttt{verify} cell records
an \texttt{engine\_tool\_gap} on the \texttt{particle} module rather than
faking a Stage-4 verdict.

\paragraph{Measured-ring optics.} The FODO twiss
(Appendix~\ref{app:fodo}) is an \emph{algorithm}-level closure on a
textbook cell. A \emph{measured}-lattice flip needs a sourced ring-optics
deck (LHC / FCC-ee / SPS) and a measured tune. This is an external-data
dependency with licensing and clean-room-risk dimensions, so it belongs
in a separate domain that ingests the facility's official optics release
--- not in the public-surface clean room.

\paragraph{Nonlinear blowout PIC.} The plasma-wakefield axis
(Appendix~\ref{app:pic}) closes the cold-linear closed form and a 1-D
\emph{linear} PIC parity. The 2-D/3-D nonlinear blowout (bubble) regime,
where the wake departs from the cold-linear shape and $E_0$ is boosted,
requires a GPU FBPIC/WarpX run with a density / drive-strength sweep ---
exactly the kind of design-grade campaign a \texttt{/micro-exp} GPU
sweep is for. It is a next-cycle item, fully named, and explicitly not
claimed by this work.

\medskip
\noindent\fbox{\parbox{0.97\linewidth}{\small\textbf{The contract.} None
of the three is framed as ``impossible with current methods'' (a g3 /
d2 violation). Each is a wet-lab / external-data / GPU dependency with a
concrete breakthrough path. The tabletop completion criterion ---
clean-room scope $\perp$ wet-lab / GPU / sourced-deck dependency --- is
the line this catalog draws, and the work inside it is complete.}}
